---
title: 'How to predict house prices?'
subtitle: 'Assignment 2 for Data Analysis 2'
author: 'Abduvosid Malikov'
geometry: "left=1cm,right=1cm,top=0.5cm,bottom=1.5cm"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


# Introduction

This is the Assignment 2 for `Data Analysis 2`  and `Coding` course. The aim of this project is  to predict house price based on its area and other variables. House price prediction can help the sellers of a house to determine the selling price of a house and can help the customer to arrange the right time to purchase a house.

# Data collection

 Data set was obtained from [Kaggle](https://www.kaggle.com/prevek18/ames-housing-dataset). It contains information from the Ames Assessor’s Office used in computing assessed values for individual residential properties sold in Ames, Iowa (IA) from 2006 to 2010.
 This data was collected by Ames Assessor’s Office, one of the cities of Iowa. This data is representative only for this state. Since this is administrative data that was collected by government office, we assume that there was no (or few) mistakes in entering data. Also, we have almost all necessary variables that matters for the house sale price (such as area, quality, rooms, etc).
 
### Data descriptives
 Variables capture the house price and other conditions such as area of a house and garage, number of rooms, availability of fireplace, pool and others. I want to use these variables to predict house price given the certain house condition (area, rooms, fireplaces).
 The data has 2930 observations. It consists of 82 columns which include 23 nominal, 23 ordinal, 14 discrete, and 20 continuous variables (and 2 additional observation identifiers, Order and Parcel ID). These variables describe the sales price of a house and its condition, such as area, quality, rooms, etc.
 As an outcome variable, we have chosen **SalePrice** (in US dollars) variable. The main parameter of interest (explanatory variable) is **Above Ground Area** (in square feet) of a house. Full summary statistics and distributions of explanatory variables are given in Appendix.

```{r setup, include=FALSE}
# install.packages("AER")
library(AER)
library(tidyverse)
require(scales)
library(lspline)
library(estimatr)
library(texreg)
library(corrplot)
# install.packages("varhandle")
library(varhandle)
library(kableExtra)
# install.packages("e1071")
library(e1071)
library(readr)
# install.packages("huxtable")
# install.packages("jtools")
library(huxtable)
library(jtools)

```


```{r,echo=F, message=FALSE, warning=FALSE}
# read clean data
df <- read.csv('../data/clean/Housing_clean.csv')


#Fireplaces (Discrete). Number of fireplaces transform to dummy:  0/1
df$HasFireplace <- ifelse(df$Fireplaces>0,1,0)
#remodeled to 0/1
df$Remodeled <- ifelse(df$Year.Remod.Add>df$Year.Built,1,0)

df$Bsmt.Qual <- df$Bsmt.Qual %>% replace_na("No Basement")
df$Bsmt.Qual[df$Bsmt.Qual ==''] <- NA
df$Bsmt.Qual.fact <- factor(df$Bsmt.Qual, order = TRUE,
                            levels = c("No Basement", "Po", "Fa", "TA", "Gd","Ex"))

df$Bsmt.Qual.unfact <- unfactor(df$Bsmt.Qual.fact)

df$Bsmt.Qual.num <- as.numeric(factor(df$Bsmt.Qual.fact,
                  levels = c("No Basement","Po", "Fa", "TA", "Gd", "Ex")))

df$MS.Zoning.num <- as.numeric(factor(df$MS.Zoning.fact))
# RM 7 RL 6 RH 5 I 4  FV 3 C 2 A 1

df$Overall.Qual.num <- unfactor(as.factor(df$Overall.Qual.fact))

df$BsmtFin.Type.1.num <- 
  as.numeric(factor(df$BsmtFin.Type.1.fact,
                    levels = c("No Basement", "Unf", "LwQ", "Rec", "BLQ", "ALQ", "GLQ")))
# GLQ 7 ALQ 6 BLQ 5 Rec 4 LwQ 3 Unf 2 No Basement 1

```
The histogram below shows the house sale prices. We can see that on average, most of the houses are priced from 100,000 to 200,000 dollars. It's skewed to the right, meaning that only few houses were priced extremely high. 


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Histogram of house sale prices"}

# Check the main parameter of interests and potential confounders:

# SalePrice
ggplot(df, aes(x = SalePrice)) +
  geom_histogram(fill='navyblue') +
  labs(x = "Sale price of houses")

```

## Data Cleaning 

### Outliers

When outcome variable SALE PRICE and parameter of interest GR LIV AREA (above ground living area square feet) were plotted, there were 5 observations with extreme values. When inspected, it was clear that two of these outliers have more than 5000 square feet area (which is extremely high) but nevertheless priced relatively appropriately. Also, three of them had Partial Sales that likely don’t represent actual
market values. Therefore, I removed all houses with more than 4000 square feet from the data set to avoid these five unusual observations (Appendix, Figure 4. 

Therefore, re-iterated research question is the following: 
**Does a house with higher above ground living area (considering it's less than 4000 square feet) have higher price from 2006 - 2010 in Ames (Iowa) ?**


### Data Type for variables
An appropriate format was given to categorical (numeric and ordinal) variables (see Appendix). 

### Missing values
Missing values in qualitative variables
As it was described in original data description, in some categorical variables, NA didn't mean missing value. Therefore, corresponding name was assigned to the observations with such values (such as "No garage" or "No garage"). 

I decided to drop columns “Alley”, “Misc.Feature”, “Fence”, “Pool.Qu” and “Fireplace.Qu” because more than 90% (48% of Fireplace.Qu) of their values were missing. 
Observations with NAs in all the garage related columns, were completed with the value “No garage” in case of nominal attribute.
I have not to used Street variable (Type of road access to property) as there are only 12 houses with Gravel and 2913 of them are Paved

# Pattern of association

With the help of scatter-plots and LOESS method, we want to check the pattern of association between *y* and each potential *x* variables. 

First, we see the plot of Sale Price and Above Ground Area since it's our main parameter of interest. The plot below shows that there is a strong linear relationship between these two variables. 


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Above Ground Area of houses"}

# Create a general function to check the pattern
chck_sp <- function(x_var, x_lab){
  ggplot( df , aes(x = x_var, y = SalePrice)) +
    geom_point() +
    geom_smooth(method="loess" , formula = y ~ x )+
    labs(y = "Sale price of houses", x = x_lab) 
}

chck_sp(df$Gr.Liv.Area, " Above ground living area square feet")  

```
 
The scatterplots that shows the pattern of association between *y variable* (Sale Price) and other explanatory variables are given in Appendix.

After we saw which kind of relationship the variables have, now we have an idea on how build a regression model and what variables to include into our model. 
 

# Comparing explanatory variables

We selected x variables that has shown strong association with y variables. Total Area, Above Ground Area, Garage Area are among them. Now we explore how these x-s are related to each other. 


```{r,echo=F, message=FALSE, warning=FALSE}

selected_var_df <- data.frame("SalePrice" = df$SalePrice, "TotalArea" = df$TotalArea, 
                              "Lot.Area" = df$Lot.Area, "HasFireplace" = df$HasFireplace, 
                              "Garage.Area" = df$Garage.Area, "Total.Bsmt.SF" =  df$Total.Bsmt.SF, 
                              "Gr.Liv.Area" = df$Gr.Liv.Area, "Bsmt.Qual" = df$Bsmt.Qual.num, 
                              "MS.Zoning" = df$MS.Zoning.num, "Overall.Qual" = df$Overall.Qual.num, 
                              "BsmtFin.Type.1" = df$BsmtFin.Type.1.num, "Year.Built" = df$Year.Built, 
                              "Year.Remod.Add" = df$Year.Remod.Add, "Garage.Yr.Blt" = df$Garage.Yr.Blt, 
                              "Garage.Cars" = df$Garage.Cars, "TotRms.AbvGrd" = df$TotRms.AbvGrd) 


# see correlation 
corTable <- round(cor(selected_var_df, use = "complete.obs"), digits = 2)
```


Table 1 shows the selected x variables and shows correlation between them.


```{r, echo = FALSE, warning = FALSE, message = FALSE}
kable(corTable, caption = "Correlation table") %>%
  kable_styling("striped", full_width = F, latex_options="scale_down") %>%
  row_spec(0, angle = 90)
```



```{r,echo=F, message=FALSE, warning=FALSE}
# Check for highly correlated values:

# Find the correlations which are higher than 0.8
id_cr <- which( abs(corTable) >= 0.8 & corTable != 1 )
pair_names <- expand.grid( variable.names(selected_var_df) , variable.names(selected_var_df) )

# Get the pairs:
high_corr <- pair_names[ id_cr , ]
high_corr <- mutate( high_corr , corr_val = corTable[ id_cr ] )

# highly correlated - confounder (leads to multicollinearity )
# This high_corr should be in Appendix as Table
```


From the correlation table, we retained the variables that has correlation less than 0.8 and removed others. The reason for this is that  highly correlated variables may be confounders that leads us to multicollinearity issue (high Standard Error). Table 2 shows highly correlated variables.

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align='center'}
kable(high_corr, caption = "Variables with high correlation") 
```

The following variables showed correlation less than 0.8, therefore I decided to retain them for further analysis and include in regression model: 

*SalePrice*: Lot.Area, HasFireplace, Garage.Area, Total.Bsmt.SF, Gr.Liv.Area, Bsmt.Qual, MS.Zoning, BsmtFin.Type.1, Year.Built, Year.Remod.Add, Garage.Yr.Blt, Garage.Cars,  TotRms.AbvGrd

I also decided to include HasFireplace variable (as a dummy variable, 1 if house has one or more fireplaces, 0 otherwise). I want to interact HasFireplace variable with Gr.Liv.Area variable (Above ground area). The reason for controlling for such interaction is that building a fireplace in the house requires additional expenditure from the owners. Therefore, owners may tend to increase house price because of the fireplace.


# Model choice

I decided to build several regression models in order to see how they perform and pick up the best one later. 


```{r, message = F, warning = F, echo = F}
# size=1, fig.height= 3, fig.align='center',
reg1 <- lm_robust(SalePrice ~ Gr.Liv.Area, data = df)
# R2 = 51

reg23 <- lm_robust(SalePrice ~ Total.Bsmt.SF + Gr.Liv.Area, data = df)
# R2 = 0.68

reg31 <- lm_robust(SalePrice ~ Gr.Liv.Area + HasFireplace + 
                     Gr.Liv.Area * HasFireplace, data = df)
# R2 = 57

reg32 <- lm_robust(SalePrice ~ Total.Bsmt.SF + Gr.Liv.Area + HasFireplace + 
                     Gr.Liv.Area * HasFireplace, data = df)
# R2 = 0.71

model_summ3 <- export_summs(reg1, reg23, reg31, reg32,
                            model.names = c("(1) First Model", "(2) Second Model",
                                            "(3) Third Model", "(4) Fourth Model"))
model_hux <- as_hux(model_summ3)
model_hux %>% set_caption("Comparing model statistics")
```


First regression model is:

**SalePrice** = Beta0 + Beta1 * **Above Ground Area**

R^2 = 51

Earlier, we have seen strong linear relationship between these variables. Even though the model fits the data pretty well, it has low R^2.


Second model is:

**SalePrice** =  Beta0 + Beta1 * **Basement Area** + Beta2 * **Above Ground Area**

R^2 = 0.68

Third model is the interaction of HasFireplace dummy variable with Above Ground Area:

Regression 3.1

**SalePrice** =  Beta0 + Beta1 * **Above Ground Area** + Beta2 * **HasFireplace** + 
                     Beta3 * **Above Ground Area** * **HasFireplace**
R2 = 57

Fourth model is:

Regression 3.2 

**SalePrice** = Beta0 + Beta1 * **Total Basement Area** +  Beta2 * **Above Ground Area** + Beta3 * **HasFireplace**                     + Beta3 * **Above Ground Area** * **HasFireplace**

R2 = 0.71


I tried to build regression model with other variables as well. These models were placed in Appendix and were not used because of low fitness (low R^2) and/or less (or not) meaningful interpretation.

Among all the models, second model was chosen for prediction of house price because it fits better than other models (high R^2, 0.68), p value of variables is significant (p < 0.05) and has meaningful interpretation. 

**SalePrice** =  Beta0 + Beta1 * **Basement Area** + Beta2 * **Above Ground Area**


From the model summary statistics table we can see the coefficients for this model. 

*Intercept (Beta0)*: -36647. Does not have meaningful interpretation

*Total.Bsmt.SF (Beta1)*: 82.30. On average, *SalePrice* is 82.30 dollars higher in the data for houses with one square feet larger *Total Basement Area * but with the same *Above Ground Living Area*.

*Gr.Liv.Area (Beta2)*: 87.63. On average, *SalePrice* is 87.63 dollars higher in the data for houses with one square feet larger *Above Ground Living Area * but with the same *Total Basement Area*.


# Prediction

Using the regression model that was selected in previous section, I tried to predict the house price. Figure 3  shows *Predicted sale prices* in x axis and *Actual sale prices* in y axis.


```{r, message = F, warning = F, echo = F, fig.cap = "Actual and predicted house prices"}
# size=1, fig.height= 3, fig.align='center',
# fig.width=7, fig.height=3, 

# y_hat-y plot - use reg23
# Predict is more general and can handle missing values...
df <- mutate( df , y_hat = predict( reg23 , df ) )

#y_hat-y plot
ggplot( data = df ) +
  geom_point (aes( x = y_hat , y = SalePrice ) ,  color="red")+
  geom_line( aes( x = SalePrice , y = SalePrice ) , color = "navyblue" , size = 1.5 )+
  labs( x = "Predicted sale prices", y = "Actual sale prices")

```

Earlier, we have seen this regression model: 

Regression 3.1

**SalePrice** =  Beta0 + Beta1 * **Above Ground Area** + Beta2 * **HasFireplace** + 
                     Beta3 * **Above Ground Area** * **HasFireplace**
                     
To understand whether fireplace interaction increases the prediction, I compared two models: model with *Above Ground Area* variable (reg23_lm) and the model where *Above Ground Area* (reg31_lm) is interacted with *HasFireplace* variable. 

To compare the prediction results of two models, I used BIC (Bayesian information criterion) and AIC (Akaike information criterion). A lower BIC means that a model is considered to be more likely to be the true model. Model with a lower AIC score will be the better-fit model.


In Table 4, regression model without interaction (reg23_lm) has shown lower BIC and AIC. Therefore, we decided to choose this model.


```{r, message = F, warning = F, echo = F, size=1, fig.align='center'}
# fig.height= 3, fig.width=3, 
reg23_lm <- lm(SalePrice ~ Total.Bsmt.SF + Gr.Liv.Area, 
               data = subset(df,complete.cases(df)))

reg31_lm <- lm(SalePrice ~ Gr.Liv.Area + HasFireplace + 
                 Gr.Liv.Area * HasFireplace, data = subset(df,complete.cases(df)))

t_BIC <- BIC(reg23_lm, reg31_lm)
t_AIC <- AIC(reg23_lm, reg31_lm)

BIC_hux <- as_hux(list(t_BIC, t_AIC))
BIC_hux  %>% set_caption("Checking model fit with BIC and AIC")
          
```


# Residual Analysis

In previous section, we predicted house prices with regression model. However, it was clear that model contained some errors - points that are either above or below the fitted line. In this section, I calculated the errors of the model and found houses with largest negative and positive errors. 

 
```{r, message = F, warning = F, echo = F}

# y_hat predicted y values from the model

# Calculate the errors of the model
df$reg23_res <- df$SalePrice - df$y_hat

# Find houses with largest negative errors
df %>% top_n( -5 , reg23_res ) %>% 
  select(PID, SalePrice, y_hat, reg23_res) %>% as_hux() %>% set_caption("Top 5 houses with largest negative errors")
  
``` 
 
Table 5 shows Actual house price (SalePrice), Predicted house price (y_hat), and the difference between them (reg23_res). For these 5 houses the model overestimated the house price, as the actual house price is cheaper  than the predicted value; in another word, these houses have lower price than average. This can be explained by the Neighborhood effect (in some Neighborhoods houses may have lower price than in others), year built (older houses may have lower price even though they have high Above Ground Area) or Overall Quality of house (has lower quality rating because of bad Utilities or bad Exterior quality).


```{r, message = F, warning = F, echo = F} 

# Find houses with largest positive errors
df %>% top_n( 5 , reg23_res ) %>% 
  select(PID, SalePrice, y_hat, reg23_res) %>% as_hux() %>% set_caption("Top 5 houses with largest positive errors")
```


Table 6 shows Actual house price (SalePrice), Predicted house price (y_hat), and the difference between them (reg23_res). For these 5 houses the model underestimated the house price, as the actual house price is more expensive  than the predicted value; in another word, these houses have higher price than average. The reasons mentioned above can be the case for this phenomenon as well. 

Overall, in Figure 3, we checked the y (Actual House Price) and y_hat (Predicted House Price) plot to examine the model fit. We can see that most scatters fall to the regression line, indicating a good fit of the model. 
 
# Robustness check

The burden for our models are external validity. The data is limited to Ames city, state Iowa from 2006 to 2010, and our prediction model was conducted using sampple data from Ames, Iowa. Applying these models to other regions, like Washington (or other US state), to commercial markets, or to time periods before 2006 or after 2010 may give the idea whether our model estimates are true in population or in generall pattern represented by the data. Additionally, the financial crisis of 2008 may had its effect to our models that made them specific to this time period and location. 

The further work may compare Ames, Iowa to other housing markets or develop a model based off housing data from a larger coverage or at a different country. This difficulty in developing these datasets is consistency in variables across regions. For example, in some countries people may prefer bigger (or smaller) houses, or houses with (or without) fireplaces.


# Appendix

### Variables

Variables describe house area, garage area, availability of fireplace, pool and other conditions. To see a full list of explanation of variables, go to [Data Documentation](http://jse.amstat.org/v19n3/decock/DataDocumentation.txt)

### Data descriptives

```{r,echo=F, message=FALSE, warning=FALSE}

get_statistics <- function(x_var){ summarise(df,
                                             n        = sum( !is.na(x_var) ),
                                             Min      = min(x_var),
                                             '1st IQR' = round(quantile(x_var, 0.25,na.rm =  T),2),
                                             Median   = median(x_var),
                                             '3rd IQR' = round(quantile(x_var,0.75, na.rm = T),2),
                                             Max      = max(x_var),
                                             Mean     = round(mean(x_var),digits = 2),
                                             Std.      = round(sd(x_var),digits = 2),
                                             Skew     = round(skewness(x_var),digits = 2)) }

s_sale_price <- get_statistics(df$SalePrice)
s_gr_liv_area <- get_statistics(df$Gr.Liv.Area)
s_basement_area <- get_statistics(df$Total.Bsmt.SF)
s_fireplaces <- get_statistics(df$HasFireplace)
s_lot_area <- get_statistics(df$Lot.Area)
s_garage_area <- get_statistics(df$Garage.Area)

SummStats <- s_sale_price %>% add_row(s_gr_liv_area) %>% add_row(s_basement_area) %>% add_row(s_fireplaces) %>% add_row(s_lot_area) %>% add_row(s_garage_area) %>% 
  add_column(c('Sale Price','Above Ground Living Area', 'Basement Area', 'Fireplaces',
               'Lot Area', 'Garage Area'))
names(SummStats)[10] <- 'Name' 

```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# kable(SummStats, caption = 'Summary Statistics of Variables') 

# SummStats %>% relocate(Name) %>% kable(caption = 'Summary Statistics of Variables ')
SummStats %>% relocate(Name) %>%  as_hux() %>% set_caption('Summary Statistics of Variables ') 

# model_summ_stat <- export_summs(SummStats)
# as_hux(model_summ_stat)

```



```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Histogram of Above Ground Area of a house"}

# Above ground living area 
ggplot(df, aes(x = Gr.Liv.Area)) +
  geom_histogram(fill='navyblue') +
  labs(x = "Above Ground Area of a house")
```

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Histogram of houses with and without fireplace"}
ggplot(df, aes(x = HasFireplace)) +
  geom_histogram(fill='navyblue') +
  labs(x = "Fireplaces")
```


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Histogram of houses with 0, 1, 2, 3 and 4 fireplaces"}

# Fireplaces
ggplot(df, aes(x = Fireplaces)) +
  geom_histogram(fill='navyblue') +
  labs(x = "Fireplaces")
```

## Data Cleaning 
clean_data.R script contains all the steps for data cleaning. Exact number of missing values in certain variables are also given as a comment

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of houses more than 4000 square feet Above Ground Area"}
df_r <- read.csv('../data/raw/Housing_raw.csv')
# Outliers
ggplot(df_r, aes(Gr.Liv.Area, SalePrice)) +
  geom_point()


```

### Data Type for variables
An appropriate format was given to categorical (numeric and ordinal) variables. In order to make it easier for further use, categorical variables were converted from character (or numeric) vector into nominal and ordinal variables (using factor and ordered factor) in clean_data.R script


## Pattern of association

Plot shows bunch of dots near 0. Probably the Sale Price and Lot Area variables are not correlated.


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price and Lot Area"}
chck_sp(df$Lot.Area, "Lot Area")
```

Ln(Sale Price) - Ln(Lot Area).  Probably linear spline, with knots at 8 and 10

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Ln(Sale Price) - Ln(Lot Area)."}

ggplot( df , aes(x = log(Lot.Area), y = log(SalePrice))) +
  geom_point() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(y = "Sale price of houses", x = "Lot Area") 
```

The more regular shape house has, the higher the price. But some Irregular houses has same price as Regular houses. 

Reg	Regular	

IR1	Slightly irregular

IR2	Moderately Irregular

IR3	Irregular

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Lot Shape"}
chck_sp(df$Lot.Shape.fact, "General shape of property")
```


Houses with 1 or more fireplaces have higher price

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Fireplace"}
chck_sp(df$HasFireplace, "Has Fireplace")
```


Almost positive linear, but some outliers at the end causing the line to curve. Seems Garage Area variable is important

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Garage Area"}
chck_sp(df$Garage.Area, "Garage Area")
```

There is a linear relationship between Sale Price and Total square feet of basement area

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Basement area"}
chck_sp(df$Total.Bsmt.SF, "Total square feet of basement area")
```

On average, houses with more garage cars have higher price
```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=6, fig.height=3, fig.cap = "Plot of Sale Price - Garage.Cars"}
chck_sp(df$Garage.Cars, "Size of garage in car capacity")
```


Near Flat/Level	and Hillside houses have higher prices.


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Land.Contour"}
chck_sp(df$Land.Contour.fact, "Land Contour")
```


There are 2922 houses with All Public *Utilities*. Since we don't have much observations with different utilities, it was decided not to check the pattern of association and not to include this variable in the model. 



Cul-de-sac and Inside lot has higher prices

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Land Configuration"}
chck_sp(df$Lot.Config.fact, " Lot configuration")
```


Houses with *Gentle slope* and *Moderate Slope* have higher prices


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=6, fig.height=3, fig.cap = "Plot of Sale Price - Land Slope"}
chck_sp(df$Land.Slope.fact, "Slope")
```


Houses from Northridge, Northridge Heights, and Stone Brook neighborhoods have relatively high prices


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Neighborhood"}
ggplot( df , aes(x = Neighborhood.fact, y = SalePrice)) +
  geom_point() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(y = "Sale price of houses", x = "Neighborhood") + 
  theme(axis.text.x = element_text(angle = 90))  
```


Houses with Normal and "Adjacent to postive off-site feature" conditions have relatively high prices

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - House Condition"}
chck_sp(df$Condition.1.fact, "Condition 1")
```


Most of the houses are Single-family Detached. Single-family Detached and Townhouse End Unit houses have relatively high prices


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Type of building"}
chck_sp(df$Bldg.Type.fact, "Type of building")
```


Most of the houses are One story and Two story. One story and Two story houses have relatively high prices


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price - Style of house"}
chck_sp(df$House.Style.fact, "Style of house")
```


Higher rating - higher sale price. But some houses has quality 6 but has same price as quality 2

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price -  House rating"}
chck_sp(df$Overall.Qual.fact, "Rating of overall material and finish of the house")
```


Most of the houses have Average (5) condition. Average conditioned houses have relatively high prices

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price -  House condition "}
chck_sp(df$Overall.Cond.fact, "Rating of the overall condition of the house")
```


The newer the house (the later is the year built) -  the higher the price. But some houses that are built before 1900 were priced realtively high. There is an upward trend


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price -  Year Built "}
chck_sp(df$Year.Built, "Year Built")
```



I also checked the pattern of association between Sale Price and Total Area of a house. TotalArea = Total Basement Area + Above Ground Area

*Total Bsmt SF* - Total square feet of basement area. (basement - below the ground floor)

*Gr Liv Area* - Above ground living area square feet


The total square footage model indicates some possible curvature (convex) which could be better interpreted with quadratic variables.


```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price -  Total Area"}
chck_sp(df$TotalArea, "Total Area of a house")
```


Total basement area and Above ground area with Sale Condition

```{r, echo = FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3, fig.cap = "Plot of Sale Price -  Total Area with Sale Condition"}
ggplot(df, aes(Total.Bsmt.SF + Gr.Liv.Area, SalePrice)) +
  geom_point(aes(color = Sale.Condition.fact, shape = Sale.Condition.fact)) + 
  labs(x = "Total basement area + Above ground area")
```


## Model choice

Regression 1.1:

SalePrice = Beta0 + Beta1 * Lot Area + Beta2 * HasFireplace

R2 = 0.27


Regression 1.2:
SalePrice = Beta0 + Beta1 * Total Area

R2 = 68

Even though this model fits the data very well, we dont take this model as TotalArea is higly correlated to SalePrice - multicollinearity issue. 


Regression 1.3:
SalePrice = Beta0 +  Beta1 * Garage Area

R2 = 42

Regression 1.4:
SalePrice = Beta0 +  Beta1 * Total square feet of basement area

R2 = 43

Regression 1.5 
SalePrice = Beta0 +  Beta1 * Basement quality

R2 = 37


Regression 1.6 
SalePrice = Beta0 +  Beta1 * MS.Zoning

R2 = 0.01

Regression 1.7 
SalePrice = Beta0 +  Beta1 * BsmtFin.Type.1

R2 = 0.11

Regression 1.8 
SalePrice = Beta0 + Beta1 * Year.Built

R2 = 0.31

Regression 1.9 
SalePrice = Beta0 + Beta1 * Year.Remod.Add

R2 = 0.29

Regression 2.0 
SalePrice = Beta0 + Beta1 * Garage.Yr.Blt

R2 = 0.28

Regresssion 2.1 
SalePrice = Beta0 + Beta1 * Garage.Cars

R2 = 0.42

Regression 2.2 
SalePrice = Beta0 + Beta1 * TotRms.AbvGrd

R2 = 0.24

```{r, message = F, warning = F, echo = F,size=1, fig.height= 3, fig.align='center'}
# Model Summary Statistics 
reg11 <- lm_robust(SalePrice ~ Lot.Area + HasFireplace, data = df)

reg12 <- lm_robust(SalePrice ~ TotalArea, data = df)

reg18 <- lm_robust(SalePrice ~ Year.Built, data = df)

reg21 <- lm_robust(SalePrice ~ Garage.Cars, data = df)

reg22 <- lm_robust(SalePrice ~ TotRms.AbvGrd, data = df)


model_summ <- export_summs(reg11, reg12, reg18, reg21, reg22 )
as_hux(model_summ)

```


Since SalePrice - TotalArea plot had a curvature, I tried to make quadratic model as well. 

Regression 2.4
SalePrice = Beta0 + Beta1 * TotalArea + Beta2 * TotalArea^2 

R2 = 0.68 - same as Regression 2.3

log - log:  ln(Price) - ln(Lot Area) piecewise linear spline

Regression 2.5 
log(SalePrice) = Alpha1 + Beta1 * log(Lot.Area)[if log(Lot.Area) < 8] + 
                (Alpha2 + Beta2 * log(Lot.Area)) * [if 8 <= log(Lot.Area) <= 10] 

R2 = 0.14


Model: Weighted linear regression, using rooms as weights.

Regression 2 

SalePrice = Above Ground Area (weights = Total Rooms Above Ground)

R2 = 0.49

Regression 3 
SalePrice = Beta0 + Beta1 * Above Ground Area + Beta2 * HasFireplace 

R2 = 55

Regression 4 
SalePrice = Beta0 + Lot.Area + Beta1 * Total Basement Area + 
                    Beta2 * Above Ground Area + Beta3 * Garage Cars + Beta4 * HasFireplace

R2 = 74

```{r, message = F, warning = F, echo = F,size=1, fig.height= 3, fig.align='center'}
reg24 <- lm_robust(SalePrice ~ TotalArea + TotalArea^2 , data = df )

reg25 <- lm_robust(log(SalePrice) ~ lspline(log(Lot.Area),c(8,10)), data = df)

reg2 <- lm_robust(SalePrice ~ Gr.Liv.Area, data = df , weights = TotRms.AbvGrd)

reg3 <- lm_robust( SalePrice ~ Gr.Liv.Area + HasFireplace, data = df )

reg4 <- lm_robust(SalePrice ~ Lot.Area + Total.Bsmt.SF + 
                    Gr.Liv.Area + Garage.Cars + HasFireplace, data = df)


model_summ2 <- export_summs(reg24, reg25, reg2, reg3, reg4)
as_hux(model_summ2)
```




